How is experience transformed into memory?

### 2018-08-28 paper/main.tex
methods edits

\subsection{Participants and Experimental Design}
Participants ($n=17$) viewed \textit{A Study in Pink}, a 50 minute episode of the BBC series, \textit{Sherlock}. Immediately upon completion of the video, participants were instructed to (verbally) recount the events in the video in their original order and in as much detail as possible. During the entire experiment, participants were in an fMRI scanner. For comprehensive details of the experimental procedures, please refer to \cite{ChenEtal17}.
%NOTE: is this sufficient or should i provide more detail?

\subsubsection{Fitting the topic model to the video text and recall transcripts}
A topic model was used to estimate the most likely mixture of topics for a given sample of text. First, the video was manually segmented into 1000 scenes, and a collection of descriptive features was manually transcribed. For each scene, we considered the following features: scene details (i.e., a sentence or two describing what happened in that scene), space (indoor or outdoor), name of all the characters in the scene, name of the character in focus, name of the character speaking, location, camera angle, music presence, and words on the screen. We concatenated the text for all of these features within each segment, creating a ``bag of words'' describing each scene. We then transformed the text descriptions into overlapping windows of 50 scene segments. For example, the first text sample comprised the text from the first 50 segments, the second comprised the text from n+1:n+51, and so on. We trained our model using these overlapping text samples with scikit-learn's (version 0.19.1) `CountVectorizer' and `LatentDirichletAllocation' classes \citep{PedrEtal11} implemented using our high-dimensional visualization/analysis software, Hypertools \citep{HeusEtal17a}.  First, the text was transformed into a vector of word counts (after removing English stopwords). This gave a word count vector for each scene in the video.  Then, the word count vectors were used to fit a topic model (topics=100, method=`batch'). We transformed the text descriptions using the model resulting in a scenes (1000) by topics (100) matrix. The scene descriptions often spanned multiple timepoints (i.e., TRs). To account for this, we expanded the video model by copying the rows of the model for as many timepoints that the scene description spanned. After this expansion, the shape of the model was the length of the duration of the video (1976 TRs).

@@

\subsubsection{Extracting events using a hidden Markov model}
The topic model timepoint-by-timepoint correlation matrices all exhibited a block-diagonal structure (with small off-diagonal values), suggesting that the models were comprised of a number of sequential `states' (or events, e.g.\ Supp. Fig.~\ref{fig:corrmats}). To capture this structure, we fit the video and each recall model using a hidden Markov model (HMM). Given a number of states or events ($k$), the HMM recovers a set of labels that segments consecutive timepoints into $k$ events ~\citep{Rabi89, BaldEtal17}. To implement this analysis, we used the Brainiak toolbox \citep{BaldEtal17, Brainiak}.

@@

\subsubsection{Matching recall events to video events}
To estimate which video event each recall event referred to, we correlated the video events model and each recall events model. This resulted in a video events (34) by recall events (8-27) correlation matrix (for each participant) which contains the similarity between each video event and each recall event.  To find the most likely video event that a given recall event referred to, we computed the argmax over the columns of this matrix.  The result was a list of video event indices for each participant. These indices are analogous to the values found in a ``recall matrix'' from a free recall list learning experiment, but represent the recall of particular events (instead of words, for example).

@@

\subsubsection{Vector field analysis}
To quantify the flow of recall from event to event, we performed a vector field analysis.  We tiled the 2D topic space (x, y: -6 to 6 by .25) with an evenly spaced grid. For each grid point, we drew a circle around the point (radius=.5). Then, we tested whether any line segments (formed by event recall transitions) passed through this area of the topic space.  For example, say that a participant transitioned from recalling event 2 to event 3. These 2 recall events correspond to 2 points in topic space, and connecting them forms a line segment. We collected all line segments that passed through a given section of topic space (collapsing across participants). To plot the average direction of the line segments (i.e., the arrows for each grid point in ~\ref{fig:trajectory}B), we converted each of them to unit vectors and then averaged. For grid points where the direction was consistent (across all participants contributing to that point), the length of the arrow approaches 1, whereas if the direction was random the length of the arrow approaches 0. Lastly, we converted each unit vector to an angle (in radians) by taking the inverse tangent of the x, y components of the vector. To test whether the distribution of angles was significantly non-uniform (i.e., displayed a preferred direction across participants), we performed a Rayleigh test on the angles ($p<.001$, FDR-corrected at $p<.05$ using Benjamani-Hochberg procedure). Arrows where the Rayleigh test was significant are displayed in color while non-significant tests are displayed in gray with lower opacity.

\subsection{fMRI analyses}
Seventeen participants watched the first 50 minutes of Episode 1 of BBC's Sherlock. The video was split into two parts of approximately equal length (946 and 1030 TRs). All data were preprocessed and transformed to 3mm MNI space as described in ~\citep{ChenEtal17}. Data were z-scored across time at every voxel. 6mm smoothing was applied.
Files are cropped so that all video-viewing data are aligned across participants, and all recall data are aligned to the scene timestamps below. The cropping includes a constant 3-TR (4.5 sec) shift to correct for hemodynamic lag.

\subsubsection{Searchlight analysis}
Our multivariate analyses were designed to capture brain regions whose timepoint-by-timepoint correlational structure mirrors the correlational structure of the video model as well as participant-specific recall topic models during video viewing. We conducted a searchlight analysis (5x5x5 voxel cube) where for each cube, we correlated the model timepoint-by-timepoint correlation matrix with the neural correlation matrix. To aggregate across participants, we Fisher's z-transformed the correlations and then averaged.  To assess significance, we recomputed this group analysis 100 times, but randomly phase shifted the model (by \# of timepoints - 1) by the same amount for each participant but different amounts for each permutation to build a null distribution of correlation values. Finally, we thresholded the group averaged correlation maps where the `real' correlation value for a given voxel exceeded the 95th percentile of the null distribution. To correct for non-linearities between the viewing time and recall time, for each participant we used dynamic time warping to temporally align the matrices. The algorithm recovers a path of coordinates that would bring the video and recall model in maximal temporal alignment. We used this path to warp the fMRI data and the recall model into temporal alignment (separately for each participant).

\subsubsection{Topic vector word clouds}
We created word clouds to visualize the themes contained in the recall events. One component of the topic model comprises a words (2117) by topics (100) matrix (R), where the rows represent the weight of a given word in each topic.  To find words that were maximally associated with a particular event vector, we computed the dot product between R and v, which gave a 1 by \# of words vector where the values represent the `activation' of each of in the event. Then, we created word clouds by extracting the top $n$ words and plotting them where the size of the word is proportional to its `activation' in the event.

In the first analysis (Fig.~\ref{fig:topics}A,B), we quantified the most and least remembered topics/words throughout the entire video by computing a weighted average over all recall events, where the weights were proportional to memory for each recall event. To measure memory for each event, for each participant we computed the correlation between the video event vector and the closest recall event vector. We then averaged these correlation values across participants. We then computed a weighted average of all video events using the correlation values as weights. Next, we computed the dot product between this weighted-average video event vector and the R matrix (described in the paragraph above) to get ``activations'' for each word. Finally, we plotted the top 200 words where the size of the word is proportional to its ``activation''. To get the least remembered topics/words, we performed the same analysis but inverted memory weights.

>>>>

\subsection{Participants and Experimental Design}
Participants ($n=17$) viewed the first 50 minutes of ``A Study in Pink'', an episode of the BBC series, \textit{Sherlock}. Immediately upon completion of the video, participants were instructed to (verbally) recount the events in the \textit{Sherlock} episode in their original order and in as much detail as possible. During the entire experiment, participants were in an fMRI scanner. For comprehensive details of the experimental procedures, please refer to \cite{ChenEtal17}.
%NOTE: is this sufficient or should i provide more detail?

\subsubsection{Fitting the topic model to the video text and recall transcripts}
A topic model was used to estimate the most likely mixture of topics for a given sample of text. First, the video was manually segmented into 1000 scenes, and a collection of descriptive features was manually transcribed. For each scene, we considered the following features: narrative details (a sentence or two describing what happened in that scene), whether the scene was indoor or outdoor, name of all the characters in the scene, name of the character in focus, name of the character speaking, location, camera angle, music presence, and text on the screen. We concatenated the text for all of these features within each segment, creating a ``bag of words'' describing each scene. We then transformed the text descriptions into overlapping windows of 50 scene segments. For example, the first text sample comprised the text from the first 50 segments, the 2nd comprised the text from n+1:n+51, and so on. We trained our model using these overlapping text samples with scikit-learn's (version 0.19.1) `CountVectorizer' and `LatentDirichletAllocation' classes \citep{PedrEtal11} implemented using our high-dimensional visualization/analysis software, Hypertools \citep{HeusEtal17a}.  First, the text was transformed into a vector of word counts (after removing English stopwords). This gave a word count vector for each scene in the video.  Then, the word count vectors were used to fit a topic model (topics=100, method=`batch'). We transformed the text descriptions using the model resulting in a scenes (1000) by topics (100) matrix. The scene descriptions often spanned multiple timepoints (i.e., TRs). To account for this, we expanded the video model by copying the rows of the model for as many timepoints that the scene description spanned. After this expansion, the shape of the model was the length of the duration of the video (1976 TRs).

@@

\subsubsection{Extracting events using a hidden Markov model}
The topic model timepoint-by-timepoint correlation matrices all exhibited a block-diagonal structure (with small off-diagonal values), suggesting that the models were comprised of a number of sequential `states' (or events, see Supp. Fig.~\ref{fig:corrmats}). To capture this structure, we fit the video and each recall model using a hidden Markov model (HMM). Given a number of states or events ($k$), the HMM recovers a set of labels that segments consecutive timepoints into $k$ events ~\citep{Rabi89, BaldEtal17}. To implement this analysis, we used the Brainiak toolbox \citep{BaldEtal17, Brainiak}.

@@

\subsubsection{Matching recall events to video events}
To estimate which video event each recall event referred to, we correlated the video events model and each recall events model. This resulted in a video events (34) by recall events (8-27) correlation matrix (for each participant) which contains the similarity between each video event and each recall event (see Supp. Fig.~\ref{matchmats}).  To find the most likely video event that a given recall event referred to, we computed the argmax over the columns of this matrix.  The result was a list of video event indices for each participant. These indices are analogous to the values found in a ``recall matrix'' from a free recall list learning experiment, but represent the recall of particular events (instead of words, for example).

@@

\subsubsection{Vector field analysis}
To quantify the flow of recall from event to event, we performed a vector field analysis.  We tiled the 2D topic space (x, y: -6 to 6 by .25) with an evenly spaced grid. For each grid point, we drew a circle around the point (radius=.5). Then, we tested whether any line segments (formed by event recall transitions) passed through this area of the topic space.  For example, say that a participant transitioned from recalling event 2 to event 3. These 2 recall events correspond to 2 points in topic space, and connecting them forms a line segment. We collected all line segments that passed through a given section of topic space (collapsing across participants). To plot the average direction of the line segments (i.e., the arrows for each grid point in ~\ref{fig:trajectory}B), we converted each of them to unit vectors and then averaged. For grid points where the direction was consistent (across all participants contributing to that point), the length of the arrow approaches 1, whereas if the direction was random the length of the arrow approaches 0. Lastly, we converted each unit vector to an angle (in radians) by taking the inverse tangent of the x, y components of the vector. To test whether the distribution of angles was significantly non-uniform (i.e., displayed a preferred direction across participants), we performed a Rayleigh test on the angles ($p<.001$, FDR-corrected at $p<.05$ using Benjamani-Hochberg procedure). Arrows where the Rayleigh test was significant are displayed in color (the darker the blue the more significant) while non-significant tests are displayed in gray with lower opacity.

\subsection{fMRI analyses}
Participants viewed and recalled the video stimulus inside an fMRI scanner. The video was split into two parts of approximately equal length (946 and 1030 TRs, $TR=1.5 seconds$). All data were preprocessed and transformed to 3mm MNI space as described in ~\citep{ChenEtal17}. Data were z-scored across time at every voxel. 6mm smoothing was applied.
Files are cropped so that all video-viewing data are aligned across participants, and all recall data are aligned to the scene timestamps below. The cropping includes a constant 3-TR (4.5 sec) shift to correct for hemodynamic lag.

\subsubsection{Searchlight analysis}
Our multivariate analyses were designed to capture brain regions whose timepoint-by-timepoint correlational structure mirrors the correlational structure of the video model as well as participant-specific recall topic models during video viewing. We conducted a searchlight analysis (5x5x5 voxel cube) where for each cube, we correlated the model timepoint-by-timepoint correlation matrix with the neural correlation matrix. To aggregate across participants, we Fisher's z-transformed the correlations and then averaged.  To assess significance, we recomputed this group analysis 100 times, but randomly phase shifted the model by the same amount for each participant but different amounts for each permutation to build a null distribution of correlation values. Finally, we thresholded the group averaged correlation maps where the `real' correlation value for a given voxel exceeded the 95th percentile of the null distribution. To correct for non-linearities between the viewing time and recall time, for each participant we used dynamic time warping to temporally align the matrices. The algorithm recovers a path of coordinates that would bring the video and recall model in maximal temporal alignment. We used this path to warp the fMRI data and the recall model into temporal alignment (separately for each participant).

\subsubsection{Topic vector word clouds}
We created word clouds to visualize the themes contained in the recall events. One component of the topic model comprises a words (2117) by topics (100) matrix (R), where the rows represent the weight of a given word in each topic.  To find words that were maximally associated with a particular event vector, we computed the dot product between R and v, which gave a 1 by \# of words vector where the values represent the ``activation'' of each of in the event. Activation is defined as the weight of a particular word in a particular event. Then, we created word clouds by extracting the top $n$ words and plotting them where the size of the word is proportional to its activation in the event.

In the first analysis (Fig.~\ref{fig:topics}A,B), we quantified the most and least remembered topics/words throughout the entire video by computing a weighted average over all recall events, where the weights were proportional to memory for each recall event. To measure memory for each event, for each participant we computed the correlation between the video event vector and the closest recall event vector. We then averaged these correlation values across participants. We then computed a weighted average of all video events using the correlation values as weights. Next, we computed the dot product between this weighted-average video event vector and the R matrix (described in the paragraph above) to get activations for each word. Finally, we plotted the top 200 words where the size of the word is proportional to its activation. To get the least remembered topics/words, we performed the same analysis but inverted memory weights.
